---
title: "Week 11: Getting data from afar: APIs"
subtitle: "PSY 525"
author: "Rick O. Gilmore"
date: "`r Sys.time()`"
css: css/gilmore.css
bibliography: bib/psu-repro.bib
csl: bib/apa.csl
output:
  ioslides_presentation:
    incremental: no
    self-contained: yes
    transition: default
    widescreen: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align = "center",
                      out.width = "700px",
                      cache = FALSE)
```

<!-- Scrolling slides -->
<!-- http://stackoverflow.com/q/38260799 -->
<style>
slides > slide { overflow: scroll; }
slides > slide:not(.nobackground):before {
  background: none;
  }
slides > slide:not(.nobackground):after {
  content: '';
  background: none;
  }
}
</style>

# Preliminaries

## Check-in

## Announcements

## Today's topics

- Tools for reproducible (remote) data gathering 

# Tools for reproducible (remote) data gathering

## What are APIs

- Application Program Interface (API)
- Talk to computer services

## Why use APIs

- Reduce human manipulation of data files
- Reduce errors
- Reduce security/privacy breaches
- Improve reproducibility

## Some useful APIs

- Box
- Google Drive
- OSF
- Databrary

## Box.com

- [`boxr` R package](https://r-box.github.io/boxr/)

## Google Drive

## OSF

## Databrary

- [`databraryapi` R package](https://github.com/PLAY-behaviorome/databraryapi)
- [`databrarypi` Python package](https://github.com/PLAY-behaviorome/databrarypi)

# Your turn

## Box

- Visit <https://psu.app.box.com/developers/console>

---


```
box_auth(client_id = "7ysxfink4ml7byn6393pwg88ub76mvkx", client_secret = "9YxPHHZUBHuqjANCpOGKjxpmE5iJMjXG")
```

# Next time...

---

- Gathering info from afar: Using APIs

# Resources

## Software

This talk was produced on `r Sys.Date()` in [RStudio](http://rstudio.com) using R Markdown.
The code and materials used to generate the slides may be found at <https://github.com/psu-psychology/psy-525-reproducible-research-2020>.
Information about the R Session that produced the code is as follows:
```{r session-info}
sessionInfo()
```

## References