---
title: "Week 13: More Python..."
subtitle: "PSY 525"
author: "Rick O. Gilmore"
date: "`r Sys.time()`"
css: css/gilmore.css
bibliography: bib/psu-repro.bib
csl: bib/apa.csl
output:
  ioslides_presentation:
    incremental: no
    self-contained: yes
    transition: default
    widescreen: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align = "center",
                      out.width = "700px",
                      cache = FALSE)
library(tidyverse)
```

<!-- Scrolling slides -->
<!-- http://stackoverflow.com/q/38260799 -->
<style>
slides > slide { overflow: scroll; }
slides > slide:not(.nobackground):before {
  background: none;
  }
slides > slide:not(.nobackground):after {
  content: '';
  background: none;
  }
}
</style>

# Preliminaries

## Today's topics

- Optional: More Python...

# More Python

## Background

- I found an excellent set of slides:  [[@Kane-2019]](https://kaneplusplus.github.io/python-for-r-programmers/).
- Rather than reinvent this excellent wheel, I suggest you work through the following:
    - [Background & Basics](https://kaneplusplus.github.io/python-for-r-programmers/part-1.html)
    - [Digging Deeper into Python Programming Constructs](https://kaneplusplus.github.io/python-for-r-programmers/part-2.html)

# Resources

## Software

This talk was produced on `r Sys.Date()` in [RStudio](http://rstudio.com) using R Markdown.
The code and materials used to generate the slides may be found at <https://github.com/psu-psychology/psy-525-reproducible-research-2020>.
Information about the R Session that produced the code is as follows:
```{r session-info}
sessionInfo()
```

## References