---
title: "Week 8: Shiny"
subtitle: "PSY 525"
author: "Rick O. Gilmore"
date: "`r Sys.time()`"
css: css/gilmore.css
bibliography: bib/psu-repro.bib
csl: bib/apa.csl
output:
  ioslides_presentation:
    incremental: no
    self-contained: yes
    transition: default
    widescreen: yes
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align = "center",
                      out.width = "900px",
                      cache = FALSE)
```

<!-- Scrolling slides -->
<!-- http://stackoverflow.com/q/38260799 -->
<style>
slides > slide { overflow: scroll; }
slides > slide:not(.nobackground):before {
  background: none;
  }
slides > slide:not(.nobackground):after {
  content: '';
  background: none;
  }
}
</style>

# Preliminaries

## Announcements

- No class next seek (Spring Break)
- HW: Complete a 1-2 page write-up describing your plans for your final project. (Commit it to the repo you're using for that work.)

## Today's topics

- Interactive web apps with Shiny

# Shiny

## Background

- RStudio framework for making interactive, web-based data applications
- Basic [tutorial](http://shiny.rstudio.com/tutorial/)
- [Gallery](http://shiny.rstudio.com/)

## Components

- User interface (UI)
- Server instructions
- Web server
- Web browser

## Template

~~~~
library(shiny)
ui <- fluidPage()
server <- function(input, output) {}
shinyApp(ui = ui, server = server)
~~~~

## Correlation demo

- **Goal**: Create an app to visualize the effects of changing the slope (m) and error (sd) in a linear relation 

$$Y \sim X * \beta + N(0,\sigma)$$

or

$$ Y \sim X_0*\beta_0 + X_1*\beta_1 + N(0,\sigma)$$ 

---

- Inputs
    + number of points
    + slope
    + error
- Outputs
    + scatter plot with fitted line
    + correlation value

---

```{r, echo=FALSE}
my.shiny.dir <- "shiny-apps/correlation-app"
shiny::shinyAppDir(appDir = my.shiny.dir,
  options=list(
    width="100%", height=700
  )
)
```

## UI input components

- Sliders for numeric input use `sliderInput()` function
- Parameters *inputId*, *label*, *min*, *max*, and *value*

~~~
sliderInput(inputId = "points",
  label = "Number of points:",
  min = 10,
  max = 200,
  value = 50)
~~~

## Sliders for slope, error

~~~
sliderInput(inputId = "slope",
  label = "Slope:",
  min = -10,
  max = 10,
  value = 1)
      
sliderInput(inputId = "error",
  label = "Error:",
  min = .0001,
  max = 5,
  value = 0.5)
~~~

## Under the hood

- Adjust the slider $\rightarrow$ the `sliderInput` function converts the slider position into a number
    - That number gets assigned to the `inputId` of the UI element
    - A list of elements with the `inputId` equal to the variable name(s) gets created: `input$points`, `input$slope`, ...
    
---

- A `server(input, output)` function takes the `ui` inputs and creates an `output`
- The UI takes the `output` and shows it on the screen

## Server function

- Syntax is `server <- function(input, output) {}`
- Use `input$points`, `input$slope`, and `input$error` to grab values from UI
- Save plot to `output$scatterPlot`

## Calculating points and plotting them

- Generate n = `input$points` as x
- Calculate y = x * `input$slope` + error
    + where error = `rnorm(input$points, sd = input$error)`
- Plot x, y

---

~~~
output$scatterPlot <- renderPlot({
    # Calculate x, y, with slope and error
    x = runif(input$points)
    
    # Vectorize x for point-wise multiplication
    y = rep(input$slope, input$points) * as.vector(x) + 
        rnorm(input$points, sd = input$error)
        
    # draw the histogram with the specified number of bins
    scatter.smooth(x = x, y = y, xlab = "x", ylab = "y")
})
~~~

## The whole shebang

```
library(shiny)

# Define UI for application that draws a scatterplot
ui <- fluidPage(
  
  # Application title
  titlePanel("Correlation Demo"),
  
  # Sidebar with a slider input for number of points 
  sidebarLayout(
    sidebarPanel(
      sliderInput(inputId = "points",
                  label = "Number of points:",
                  min = 10,
                  max = 200,
                  value = 50),
      sliderInput(inputId = "slope",
                  label = "Slope:",
                  min = -10,
                  max = 10,
                  value = 1),
      sliderInput(inputId = "error",
                  label = "Error:",
                  min = .0001,
                  max = 5,
                  value = 0.5)
    ),
    
    # Show a plot of the generated distribution
    mainPanel(
      plotOutput(outputId = "scatterPlot")
    )
  )
)

# Define server logic
server <- function(input, output) {
  
  output$scatterPlot <- renderPlot({
    
    # Calculate x, y, with slope and error
    x = runif(input$points)
    y = rep(input$slope, input$points) * as.vector(x) + rnorm(input$points, sd = input$error)
    
    # draw the plot
    scatter.smooth(x = x, y = y, xlab = "x", ylab = "y")
  })
}
```

## Why do this?

- Fun, informative way to **simulate** before you run your study
- Strong hypothesis generation, prediction
- Interactive report
- Interactive publication!

## Let's explore

- <http://wordbank.stanford.edu>
- <https://www.rstudio.com/products/shiny/shiny-user-showcase/>
- <http://shinyapps.org/>
- <https://radiant-rstats.github.io/docs/>
- <https://www.showmeshiny.com/>

## Learning more

- Shiny cheatsheet from RStudio. <https://www.rstudio.com/wp-content/uploads/2016/01/shiny-cheatsheet.pdf>

# Your turn

## Complete a series of "quests"

- Clone the repo at <https://github.com/psu-psychology/shiny-resources>
- Render `shiny-templates/shiny-work.Rmd` to see what the various examples do using `rmarkdown::render("shiny-templates/shiny-work.Rmd")`.
- Try the quests!
- Have fun

# Next time...

---

- Python!
