---
title: "Interact with the U.S. Census API"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

# Purpose

This document describes how to interact with U.S. Census data.
This allows researchers to collect data relevant to their research questions.

# Prerequisites

- You must have a Census API key.

Visit <https://api.census.gov/data/key_signup.html> to apply for an API key.
The key will be delivered to you via email.

# Installation

## Package dependencies

We'll use the [`tidycensus` package](https://walkerke.github.io/tidycensus/).
Most of the steps in this brief how_to come from <https://walkerke.github.io/tidycensus/articles/basic-usage.html>.

```{r install-packages}
if (!require(tidycensus)){
  install.packages("tidycensus")
}

# Load tidyverse package for pipe `%>%` operator
library(tidyverse)
```

## Store credentials

Run the following command at the R console, inserting the API key you received from the Census.

```{r, eval=FALSE}
tidycensus::census_api_key("YOUR API KEY GOES HERE")
```

To install for future sessions, run the following command:

```{r, eval=FALSE}
tidycensus::census_api_key("YOUR API KEY GOES HERE", install = TRUE)
```

This saves the Census API key in your `.Renviron` variable.

# Testing the package

The [tutorial](https://walkerke.github.io/tidycensus/articles/basic-usage.html) suggests the following command to test the API:

```{r}
med_rent_1990 <- tidycensus::get_decennial(geography = "state", 
                                         variables = "H043A001", year = 1990)
head(med_rent_1990)
```

This downloads data about median rents by state from the 1990 Census.

```{r}
med_rent_1990 %>%
  ggplot(aes(x = value, y = reorder(NAME, value))) + 
  geom_point()
```

Here's another test of the package.
Let's look at household income by county in Pennsylvania from the 2018 American Community Survey (ACS).

```{r}
pa <- tidycensus::get_acs(geography = "county", 
              variables = c(medincome = "B19013_001"), 
              state = "PA", 
              year = 2018)

pa
```

With a bit of cleaning, we can make a nice plot.

```{r}
pa %>%
  mutate(NAME = gsub(" County, Pennsylvania", "", NAME)) %>%
  ggplot(aes(x = estimate, y = reorder(NAME, estimate))) +
  geom_errorbarh(aes(xmin = estimate - moe, xmax = estimate + moe)) +
  geom_point(color = "red", size =1) +
  labs(title = "Household income by county in Pennsylvania",
       subtitle = "2013-2017 American Community Survey",
       y = "",
       x = "ACS estimate (bars represent margin of error)")
```

