---
title: "Interact with the Google Drive API"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

# Purpose

This document describes how to interact with the Google Drive API.
This enables users to automate file uploading, downloading, and sharing.
The [documentation site](https://googledrive.tidyverse.org/) is good.
This brief how-to borrows from it.

# Prerequisites

- You must have a personal Google account or have your Google apps account via Penn State set-up.

# Installation

## Package dependencies

We'll use the [`googledrive`](https://googledrive.tidyverse.org/) package.

```{r install-packages}
if (!require(googledrive)){
  install.packages("googledrive")
}
```

# Test the connection to Google Drive

## List files

List files in your Google Drive.

```{r}
googledrive::drive_find(n_max = 5)
```

This will initiate the process of authenticating to your Google Drive account.
Depending on which Google Drive account you choose (e.g., your personal one or a PSU one associated with your access ID like `rog1@psu.edu`) there will be different sets of steps.

When you see "Authentication complete. Please close this page and return to R.", you have finished the authentication process.
If you have documents in your Google Drive under the account you chose, you will see them listed in the console.

## Upload files

Let's create a test file and then upload it.

```{r create-test-data}
df <- data.frame(name = c("Tom", "Dick", "Harriet"),
                 age = c(10, 15, 20),
                 alive = c(FALSE, TRUE, TRUE))

readr::write_csv(df, "tmp.csv")

# Upload tmp.csv as test-data and make it a spreadsheet
(test_data <- googledrive::drive_upload(
  media = "tmp.csv",
  name = "my-525-test-data",
  type = "spreadsheet"
))
```

We now confirm that Google drive knows about it.

```{r find-test-data}
googledrive::drive_find(pattern = "my-525-test-data")
```

It does, so let's download it again.

```{r}
new_test <- googledrive::drive_download("my-525-test-data")
```
This downloads the file as `.xlsx`.

# Sheets-specific commands

## Package installation

You will want to install the [`googlesheets4` package](https://googlesheets4.tidyverse.org/) to interact with Google sheets.

```{r}
if (!require(googlesheets4)){
  install.packages("googlesheets4")
}
```

## Reading sheets

```{r}
gapm_data <- googlesheets4::read_sheet("1U6Cf_qEOhiR9AZqTqS3mbMF3zt2db48ZP5v3rkrAEJY")
```

The first time you call this command, the package will ask to confirm your authentication details.

Now, we can visualize the data.

```{r}
str(gapm_data)
```

```{r}
library(ggplot2)
gapm_data %>%
  ggplot(.) +
  aes(x = lifeExp, y = country) +
  geom_violin()
```

