---
title: "Interact with KoboToolbox API"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

# Purpose

This document describes how to interact with the KoboToolbox API.
KoboToolbox is a tool for creating web-based surveys.
The surveys can also be run on an Android device offline, and the results pushed to the server when the device is back online.

# Prerequisites

- An account on KoboToolbox.

Visit <https://www.kobotoolbox.org/> and click on the 'Sign Up' menu item.

Then press the 'Create An Account' button on the right side of the screen if you are a researcher.

# Installation

## Package dependencies

We'll use the [`koboloadeR` package](https://unhcr.github.io/koboloadeR/).
It's still in beta stage, so we have to install it via the `devtools` package from the package's GitHub repo.

```{r}
if (!require(devtools)) {
  install.packages("devtools")
}
devtools::install_github("unhcr/koboloadeR", ref = "gh-pages") 
```

