---
title: "Python and R Markdown"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

# Purpose

This document describes how to use Python in an R Markdown document.

This tutorial relies upon and extends this site: <https://rstudio.github.io/reticulate/articles/r_markdown.html>

# Prerequisites

## Install Python

You want to use Python 3.
The latest version as of `r Sys.Date()` is 3.8.2, but I recommend using the most recent stable version (3.7.6).
Python 2 is no longer under active development.
There is a still-smouldering flame war between the Python 2 and Python 3 communities.
Remember my adage--avoid flame wars.
Go with Python 3.

### Mac OS X

If you own your own machine, I strongly recommend using the Homebrew package manager. 
Follow the instructions [here](https://brew.sh/) to install Homebrew.
Then, you can enter `brew install python3` in Terminal.app to install Python 3.

It's fine if you want to install Python 3 directly from the Python [site](https://www.python.org/downloads/mac-osx/) or use [Anaconda](https://www.anaconda.com/) or another tool.
My main beef with the Conda package manager is that it seems to support fewer Python packages than the alternatives.
So, I recommend installing Python via Homebrew and using `pip` to install additional packages.

Check that this worked by entering `python3 --version` to check the installed Python verson.
You might also want to run `which python3` and note the directory where the Python 3 source code is located.

You'll want to use that as a variable in a future step.

For example, here's a code chunk that does this:

```{r}
path_2_python3 <- system("which python3")
if (path_2_python3 == "") message("Python3 not found.")
```

You may find [this site](https://www.digitalocean.com/community/tutorials/how-to-install-python-3-and-set-up-a-local-programming-environment-on-macos) helpful if you run into trouble.

### Windows

[Here](https://www.python.org/downloads/windows/) is the official Python site for Windows.
[Anaconda](https://www.anaconda.com/) is also an option, but see my caveat above.

I was able to install Python on one of my personal PCs using the standard download procedure from <https://www.python.org/downloads/windows/>.

### An alternative

It appears that the `reticulate` package will ask if you want to install the `mini-conda` package manager if it does not find a Python distribution.
It's fine to use this if it suits your needs.

## Install the R `reticulate` package

Enter `install.packages("reticulate")` from your R console.

# Installing Python packages

Like R, bare bones Python is useful but limited.
You'll want to install packages to extend its functionality.

## Mac OS

Open a terminal, and enter the following commands:

```
pip3 install pandas
pip3 install numpy
pip3 install matplotlib
```

These three packages give you much-needed additional functionality.

**Note:**`pip3` is the Python 3 version of `pip`. Depending on your specific set-up, it's possible that you could use `pip install` here, but the `pip3 install` command ensures that you install the right versions of the packages.

If you do not have admin privileges, you may need to use these commands to install the packages with the `--user` flag to limit the installation to your own account.

```
pip3 install pandas --user
pip3 install numpy --user
pip3 install matplotlib --user
```

## Windows

Installation on Windows is essentially the same.
Open the Command Prompt application, then enter these commands:

```
pip3 install pandas
pip3 install numpy
pip3 install matplotlib
```

or as needed

```
pip3 install pandas --user
pip3 install numpy --user
pip3 install matplotlib --user
```

# Test set-up

Load the `reticulate` R package and specify path to the Python version we want to use.
Then check to make sure that `reticulate` can talk to Python 3.
(Note that this next chunk is an R chunk).

```{r}
library(reticulate)
if (reticulate::py_available()) message("Python 3 found.")
```

## Install a few packages

Let's check to see if `pandas` is installed.
Pandas is the Python world's equivalent of `tidyverse` although in saying so I'm sure I'm offending someone.

```{r python-setup}
if (reticulate::py_module_available("pandas")) message("'pandas' found.")
```

We'll also check to see if `matplotlib` is installed.
This is a core plotting library.

```{r}
if (reticulate::py_module_available("matplotlib")) message("'matplotlib' found.")
```

If these are *not* installed, then I suggest you install the packages outside of R, as described above.

# Working example

We import the `pandas` package and the `csv/zoo.csv` dataset as a pandas data frame called `critters`.

```{python}
import pandas
critters = pandas.read_csv("csv/zoo.csv")
```

Then, we print the head of the dataset.

```{python}
# default is 5
critters.head(n=3)
```

Note the syntax.
When we created `critters`, it created a pandas data frame.
Pandas data frames allow a number of methods (functions) to be applied to them, `head()` is one of them.
So, to apply the `head()` function to `critters`, we put method/function call at the end.

Here's the full `critters` data set.

```{python}
print(critters)
```

```{python}
pandas.crosstab(critters['animal'], columns = ['animal'])
```
