---
title: "Python and R Markdown"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

# Purpose

This document describes how to use Python in an R Markdown document.

This tutorial relies upon and extends this site: <https://rstudio.github.io/reticulate/articles/r_markdown.html>

# Prerequisites

## Install Python

Follow the instructions at <https://psu-psychology.github.io/psy-525-reproducible-research-2020/how_to/install-python.html> to install Python on your local machine.

## Confirm that your R Studio session can see Python3

The following works on Mac OS.
I need to add a more generic test that works on Windows, too.

```{r}
path_2_python3 <- system("which python3")
if (path_2_python3 == "") message("Python3 not found.")
```

You may find [this site](https://www.digitalocean.com/community/tutorials/how-to-install-python-3-and-set-up-a-local-programming-environment-on-macos) helpful if you run into trouble.

## An alternative

It appears that the R `reticulate` package will ask if you want to install the `mini-conda` package manager if it does not find a Python distribution.
It's fine to use this if it suits your needs.
I have not tested it yet.

# Install the R `reticulate` package

Enter `install.packages("reticulate")` from your R console.

# Install Python packages

Like R, bare bones Python is useful but limited.
You'll want to install packages to extend its functionality.

## Mac OS

Open a terminal, and enter the following commands:

```
pip3 install pandas
pip3 install numpy
pip3 install matplotlib
```

These three packages give you much-needed additional functionality.

**Note:**`pip3` is the Python 3 version of `pip`. Depending on your specific set-up, it's possible that you could use `pip install` here, but the `pip3 install` command ensures that you install the right versions of the packages.

If you do not have admin privileges, you may need to use these commands to install the packages with the `--user` flag to limit the installation to your own account.

```
pip3 install pandas --user
pip3 install numpy --user
pip3 install matplotlib --user
```

## Windows

Installation on Windows is essentially the same.
Open the Command Prompt or PowerShell application, then enter these commands:

```
pip3 install pandas
pip3 install numpy
pip3 install matplotlib
```

or as needed

```
pip3 install pandas --user
pip3 install numpy --user
pip3 install matplotlib --user
```

# Test set-up

Load the `reticulate` R package and specify path to the Python version we want to use.
Then check to make sure that `reticulate` can talk to Python 3.
(Note that this next chunk is an R chunk).

```{r}
library(reticulate)
if (reticulate::py_available()) message("Python 3 found.")
```

## Install a few packages

Let's check to see if `pandas` is installed.
Pandas is the Python world's equivalent of `tidyverse` although in saying so I'm sure I'm offending someone.

```{r python-setup}
if (reticulate::py_module_available("pandas")) message("'pandas' found.")
```

We'll also check to see if `matplotlib` is installed.
This is a core plotting library.

```{r}
if (reticulate::py_module_available("matplotlib")) message("'matplotlib' found.")
```

If these are *not* installed, then I suggest you install the packages outside of R, as described above.

# Working example

We import the `pandas` package and the `csv/zoo.csv` dataset as a pandas data frame called `critters`.

```{python}
import pandas
critters = pandas.read_csv("csv/zoo.csv")
```

Then, we print the head of the dataset.

```{python}
# default is 5
critters.head(n=3)
```

Note the syntax.
When we created `critters`, it created a pandas data frame.
Pandas data frames allow a number of methods (functions) to be applied to them, `head()` is one of them.
So, to apply the `head()` function to `critters`, we put method/function call at the end.

Here's the full `critters` data set.

```{python}
print(critters)
```

```{python}
pandas.crosstab(critters['animal'], columns = ['animal'])
```
