---
title: "Making Reports with Parameters"
subtitle: "PSY 525"
author: "Rick Gilmore"
date: "`r Sys.time()`"
output: 
  html_document:
    self_contained: true
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      out.width = "400px",
                      fig.align = "center")
```

# Purpose

This document describes how to create "parameterizable" reports.

You could use this to create one R Markdown document that summarizes data from single individuals across a study.

# Preparatory work

For our example use case, we need a set of data files for each participant.
The following code generates a set of data files and stores them in the `csv/` directory.
We've set `eval=FALSE` since we only need to run this once.

```{r, warning=FALSE, message=FALSE, echo=TRUE, eval=FALSE}
n_subs <- 5

letts <- letters[1:3]
nums <- 1:3
stims <- c(letts, nums)

n_blocks_run <- 10
n_runs_expt <- 5

make_conds <- function(stims, n_blocks_run, 
                       n_runs_expt) {
  x <- rep(sample(stims, length(stims)), n_blocks_run)
  rep(x, n_runs_expt)
}

simulate_rt <- function(df, mu = 300, mu_sd = 20, 
                        err_sd = 20, lett_eff = 40) {
  n_trials <- dim(df)[1]
  sub_mu <- rnorm(1, mu, mu_sd)
  mus <- rep(sub_mu, n_trials)
  errs <- rnorm(n_trials, mean = 0, sd = err_sd)
  dplyr::mutate(df, rt = mus + ifelse(type == "letters", lett_eff, 0) + errs)
}

make_data <- function(p_id) {
  stims <- make_conds(stims, n_blocks_run, n_runs_expt)
  df <- data.frame(p_id = p_id,
                   stim = stims)
  df <- dplyr::mutate(df, type = ifelse(stim %in% letters, "letter", "number"))
  df
}

save_data <- function(df, data_dir = "csv") {
  if (!dir.exists(data_dir)) dir.create(data_dir)
  fn <- paste0("p_", unique(df$p_id), ".csv")
  fn_full <- file.path(data_dir, fn)
  message("Saving data to '", fn_full, "'.")
  readr::write_csv(df, fn_full)
}

make_save_data <- function(p_id) {
  d <- make_data(p_id)
  d <- simulate_rt(d)
  save_data(d)
}

purrr::map(1:n_subs, make_save_data)
```

# Procedure

## Create a report template.

Your template must have a `params:` value in the YAML header.
This is the means by which we'll send the document parameters.
It's good practice to give a default value for the parameter.
That let's us knit (render) the report while we are tweaking it.
We can always change the parameter value(s) when we render the report from the console.

Here is an example from the `report-template.Rmd` file.

````yaml
---
title: "Template Parameterizable Report"
author: "Rick Gilmore"
date: "`r Sys.time()`"
output: html_document
params:
  data_file_path: "csv/p_01.csv"
---
````

## Add chunk to import

Add a chunk to the template to import your data.
Use `params$data_file_path` as the file parameter to import.

````markdown
`r ''````{r import-data}
these_data <- readr::read_csv(file = params$data_file_path)
```
````

## Add other chunks 

Add chunks to to visualize or analyze your data.

This is the chunk from the report template that plots a histogram for a single participant.

````markdown
`r ''````{r plot-hist}
ggplot2::ggplot(these_data) +
  ggplot2::aes(x = rt, fill = type) +
  ggplot2::geom_histogram()
```
````

## Create report-rendering functions

I've saved the following under `R/report_rendering_helpers.R`.

### Render single report

This function renders a single report 

```
render_my_report <- function(fn, out_dir = "html") {
  if (!file.exists(fn)) {
    stop("File '", fn, "' not found.")
  }
  if (!dir.exists(out_dir)) {
    dir.create(out_dir)
  }
  
  # Create output file name
  report_fn <- paste0(fn, ".html")
  
  rmarkdown::render(input = "report-template.Rmd", 
                    output_file = report_fn,
                    output_dir = "html", 
                    params = list(data_file_path = fn))
}
```

### Render list of reports

This function runs the report across the list of input files using the `purrr:map()` function.
This style of programming avoids ugly "for" loops, and is super-efficient, but is a bit advanced for us at this juncture.

```
render_all_reports <- function(data_dir = "csv") {
  if (!dir.exists(data_dir)) {
    stop(paste0("Data directory '", data_dir, "' not found."))
  }
  
  fl <- list.files(data_dir, pattern = "p_[1-9]", full.names = TRUE)
  purrr::map(fl, render_my_report)
}
```

## Render the reports

"Source" your `report_rendering_helpers.R` to bring the helper functions into your current environment.
Then, run the `render_all_reports()` command from the console.

*Note*: This code assumes that your R Markdown template is in your current directory, and that the `R/`, `csv/`, and `html/` directories are available in your current directory.
Since this tutorial is in a sub-directory of the course website, I changed my working directory to `how_to/` before running the code.
If you want to try this yourself, I recommend copying the relevant files to a clean project so you don't have to change your working directory.

*Note*: Do this sort of "meta-programming" with R Markdown from R scripts and the console.
R Markdown doesn't do well with rendering other R Markdown documents.
It's like that scene in ["Being John Malkovich"](https://en.wikipedia.org/wiki/Being_John_Malkovich).

## View the reports

[html/p_1.csv.html](html/p_1.csv.html)

[html/p_2.csv.html](html/p_2.csv.html)

[html/p_3.csv.html](html/p_3.csv.html)

[html/p_4.csv.html](html/p_4.csv.html)

[html/p_5.csv.html](html/p_5.csv.html)
